---
title: "Figure 3"
author: "YueanZhao"
output: html_document
---

```{r}
scRNAsubT8 <- readRDS("rds/scRNAsubT8_anno_TCRpercent_Cycling.NoCD4.rds")
DefaultAssay(scRNAsubT8) <- "RNA"

scRNAsubT8$cell_type_main <- factor(scRNAsubT8$cell_type_main, levels =order_T8_main)
scRNAsubT8$TCR_clonetype <- factor(scRNAsubT8$TCR_clonetype, levels =order_TCR_clonetype)
scRNAsubT8$disease <- factor(scRNAsubT8$disease,levels = order_disease)
scRNAsubT8$patient <- factor(scRNAsubT8$patient,levels = order_patientID)

```

# fig3a: clonetype expansion dimplot

```{r}
scRNA <- scRNAsubT8[,scRNAsubT8$patient != "C2"]
md <- cbind(Embeddings(scRNA[["umap"]]),scRNA[["TCR_clonetype"]],scRNA[["disease"]])

p = ggplot(md[!is.na(md$TCR_clonetype),], 
           aes_string(x = "UMAP_1", y = "UMAP_2", color = "TCR_clonetype"))+
  geom_point(data = md[is.na(md$TCR_clonetype) & md$disease == "Pneumonia",], color = "lightgrey", size = 1)  + 
  geom_point(data = md[!is.na(md$TCR_clonetype) & md$disease == "Pneumonia",],shape = 16,size = 1,alpha = 0.8)  +
  scale_color_manual(values = color_TCR_clonetype_disease)+
  labs(colour = "Pneumonia")+
  theme_cowplot()+
  theme_blank+NoLegend()

ggsave("fig3/3a_DimPlot_T8_clonetype_disease.png",p, width = 5, height = 5)
```


```{r}
p = ggplot(md[!is.na(md$TCR_clonetype),], 
           aes_string(x = "UMAP_1", y = "UMAP_2", color = "TCR_clonetype"))+
  geom_point(data = md[is.na(md$TCR_clonetype) & md$disease == "Ctrl",], color = "lightgrey", size = 1)  +
  geom_point(data = md[!is.na(md$TCR_clonetype) & md$disease == "Ctrl",],
             aes_string(x = "UMAP_1", y = "UMAP_2", color = "TCR_clonetype"),
             shape = 16,size = 1,alpha = 0.8) +
  scale_color_manual(values = color_TCR_clonetype_ctrl)+
  labs(colour = "ctrl")+
  theme_cowplot()+
  theme_blank+NoLegend()
ggsave("fig3/3a_DimPlot_T8_clonetype_ctrl.png",p, width = 5, height = 5)
```

```{r}
scRNA <- scRNAsubT8[,scRNAsubT8$patient != "C2"]
md <- cbind(Embeddings(scRNA[["umap"]]),scRNA[["TCR_clonetype"]],scRNA[["disease"]])

p = ggplot(md[!is.na(md$TCR_clonetype),], 
           aes_string(x = "UMAP_1", y = "UMAP_2", color = "TCR_clonetype"))+
  geom_point(data = md[is.na(md$TCR_clonetype),], color = "lightgrey", size = 1)  + 
  geom_point(data = md[!is.na(md$TCR_clonetype) & md$disease == "Pneumonia",],shape = 16,size = 1,alpha = 0.8)  +
  scale_color_manual(values = color_TCR_clonetype_disease)+
  labs(colour = "Pneumonia")+ new_scale("color") + 
  geom_point(data = md[!is.na(md$TCR_clonetype) & md$disease == "Ctrl",],
             aes_string(x = "UMAP_1", y = "UMAP_2", color = "TCR_clonetype"),
             shape = 16,size = 1,alpha = 0.8) +
  scale_color_manual(values = color_TCR_clonetype_ctrl)+
  labs(colour = "ctrl")+
  theme_cowplot()+
  theme_blank+NoLegend()

ggsave("fig3/3a_DimPlot_T8_clonetype_merge.png",p, width = 5, height = 5)
```

```{r}
p = ggplot(md[!is.na(md$TCR_clonetype),], 
           aes_string(x = "UMAP_1", y = "UMAP_2", color = "TCR_clonetype"))+
  geom_point(data = md[is.na(md$TCR_clonetype) & md$disease == "Pneumonia",], color = "lightgrey", size = 1)  + 
  geom_point(data = md[!is.na(md$TCR_clonetype) & md$disease == "Pneumonia",],shape = 16,size = 1,alpha = 0.8)  +
  scale_color_manual(values = color_TCR_clonetype_disease)+
  labs(colour = "Pneumonia")+new_scale("color")+
  geom_point(data = md[!is.na(md$TCR_clonetype) & md$disease == "Ctrl",],
             aes_string(x = "UMAP_1", y = "UMAP_2", color = "TCR_clonetype"),
             shape = 16,size = 1,alpha = 0.8) +
  scale_color_manual(values = color_TCR_clonetype_ctrl)+
  labs(colour = "Ctrl")+
  theme_cowplot()+
  theme_blank
l=as_ggplot(get_legend(p)) 
ggsave('fig3/3a_DimPlot_T8_clonetype_legend.pdf', l)
```

# fig3b: clonetype expansion
```{r}
scRNA <- scRNAsubT8[,scRNAsubT8$patient != "C2"]
clusters <- data.frame(table(scRNA$TCR_clonetype, scRNA$disease,scRNA$cell_type_main))
colnames(clusters)[1:3]<- c("TCR_clonetype","group","cell_type")
clusters <- clusters[order(clusters$cell_type,clusters$group,clusters$TCR_clonetype),]

p = ggplot(clusters) + 
  geom_bar(data = clusters[clusters$group == "Pneumonia",],
           aes(x =cell_type, y= Freq, fill = TCR_clonetype, group = group),
           position ="stack",stat = "identity",width = 0.4,just = 1) + 
  scale_fill_manual(values=c("#FFC30F", "#FF5744", "#C70039", "#900C3F", "#581845"),
                    breaks = rev(order_TCR_clonetype))+
  labs(fill = "Pneumonia")+new_scale("fill")+
  geom_bar(data = clusters[clusters$group == "Ctrl",],
           aes(x =cell_type, y= Freq, fill = TCR_clonetype, group = group),
           position ="stack",stat = "identity",width = 0.4,just = 0) + 
  scale_fill_manual(values=c("#C6DBEF", "#9ECAE1", "#6BAED6", "#4292C6", "#2171B5"),
                    breaks = rev(order_TCR_clonetype))+
  labs(fill = "Ctrl")+
  labs(x=NULL,y = 'Number of Cells')+
  theme_cowplot()+
  theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1,size = rel(1)),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14))+
  scale_y_continuous(expand = c(0,0))+
  coord_cartesian(ylim = c(0, 5100))

ggsave("fig3/3b_T8_clonetype_stackbar.pdf",p,width = 8, height = 5)
```

# fig3c tcr transition index
```{r}
meta <- scRNAsubT8@meta.data
meta$Cell_Name <- rownames(meta)
meta$clone.id <- meta$TCR_ID
meta$majorCluster <- meta$cell_type_main
meta$loc <- meta$sample_type
meta <- subset(meta, meta$clone.id != "NA")
tic("Startrac.run")
out <- Startrac.run(meta, proj="T8",verbose=F)
toc()

df = out@pIndex.tran
df <- subset(df, df$aid == "T8")
rownames(df) <- df$majorCluster
df$aid <- NULL
df$majorCluster <- NULL
df <- myfill(df)
p = Heatmap(df,name = "pIndex.tran", column_names_rot = 45,
        col=colorRampPalette(rev(brewer.pal(n = 7, name ="RdBu")))(100),
        heatmap_legend_param = list(labels_gp = gpar(font = 16),direction = "vertical"),
        width = unit(2.5, "in"), height = unit(2.5, "in"),
        column_names_gp = gpar(fontsize = 14),row_names_gp = gpar(fontsize = 14))
pdf("fig3/3c_transition_index.pdf",width = 5,height = 4)
draw(p)
dev.off()
```

# fig3d: TCR heatmap (group:disease) 
## TCR top10
```{r}
TCR_top10 = as.data.frame(table(scRNAsubT8[,scRNAsubT8$bchain_ID_sample <= 10 & 
                                          scRNAsubT8$sample_type %in% "Lung" & 
                                          scRNAsubT8$disease %in% "Pneumonia"]$TCR_ID))$Var1

scRNA_top10 <- scRNAsubT8[,scRNAsubT8$TCR_ID %in% TCR_top10]
```

## disease heatmap
### matrix
```{r}
mat <- scRNAsubT8@meta.data[,c("sample_type","TCR_ID","TCR_percent","group")]
mat <- mat %>% drop_na()
mat <- mat[!duplicated(mat),]
mat <- subset(mat, mat$TCR_ID %in% TCR_top10 )
mat$TCR_percent <- mat$TCR_percent*100
```

```{r}
mat1 <- mat %>% mutate(log_percent = log10(TCR_percent+1))
mat1 <- dplyr::select(mat1,-3)
```

```{r}
mat_cycling <- mat1
```

```{r}
mat1 <- dcast(mat1,TCR_ID~sample_type, value.var = "log_percent")
rownames(mat1) <- mat1[,1]
mat1 <- mat1[,-1]  
```

```{r}
myfill <- function(df){
  df2 <- df %>%
    mutate(across(where(is.character),function(x){ replace(x,is.na(x),"0")})) %>%
    mutate(across(where(is.numeric),function(x) {replace(x,is.na(x),0)}))
  return(df2)
}
```

```{r}
mat1 <- myfill(mat1)
```

### annotation
```{r}
df <- scRNAsubT8@meta.data[,c("TCR_ID","patient")]%>% drop_na()
df <- df[!duplicated(df),]
df <- subset(df, df$TCR_ID %in% TCR_top10)%>% dplyr::arrange(TCR_ID)
rownames(df) <- df[,1]
df$patient <- factor(df$patient,levels = c("P2","P3","P4","P5","P6","P7","P8","P9","P10","P11","P12"))

color_cluster <- c("#ED7216","#BFDB38","#1F8A70")

la <- rowAnnotation(cluster = anno_block(width = unit(5, "mm"),
                                          gp = gpar(fill = color_cluster,col=NA),
                                          labels =c("pattern 1a", "pattern 1b", "pattern 2"), 
                                          labels_gp = gpar(col = "white", 
                                                           fontsize = 12,
                                                           fontface = "bold")),
                    patientID = df$patient, 
                    col = list(patientID = color_patientID),
                    simple_anno_size = unit(4, "mm"),
                    annotation_legend_param = list(patientID = list(labels_gp = gpar(fontsize = 12),
                                                                    title_gp = gpar(fontsize = 13,
                                                                                    face = "bold"))))
```

### cycling label
```{r}
cycling <- scRNAsubT8@meta.data[,c("TCR_ID","group","sample_type","cell_type_T8")]%>% drop_na()
cycling <- subset(cycling, cycling$TCR_ID %in% TCR_top10 & 
                    cell_type_T8 %in% "Cycling" )%>% dplyr::arrange(TCR_ID)
cycling <- cycling[!duplicated(cycling),]
cycling$cycling <- "*"

cycling <- tidyr::unite(cycling,"matchID","TCR_ID","sample_type",sep="_",remove=FALSE)
mat_cycling <- tidyr::unite(mat_cycling,"matchID","TCR_ID","sample_type",sep="_",remove=FALSE)

mat_cycling$cycling <- cycling$cycling[match(mat_cycling$matchID, cycling$matchID)]

mat_cycling <- dcast(mat_cycling,TCR_ID~sample_type, value.var = "cycling")
rownames(mat_cycling) <- mat_cycling[,1]
mat_cycling <- mat_cycling[,-1]  
```

### heatmap
```{r}
set.seed(200)
p = Heatmap(mat1, name = "lg(TCR%+1)", 
                         cluster_rows = T, cluster_columns = FALSE,na_col = 'grey',
                         row_km = 3,
                         column_names_gp = gpar(fontsize = 14),
                         column_title = "TCR abundance heatmap",
                         left_annotation = la,
                         show_row_names = F,clustering_method_rows = "complete",
                         col=colorRampPalette(rev(brewer.pal(n = 7, name ="RdYlBu")))(100),
                         heatmap_legend_param = list(labels_gp = gpar(fontsize = 12),
                                                     direction = "vertical",
                                                     title_gp = gpar(fontsize = 13,face = "bold")),
                         cell_fun = function(j, i, x, y, width, height, fill) {
                           if(mat_cycling[i, j] %in% "*")
                             grid.text(sprintf(mat_cycling[i, j]), x, y, gp = gpar(fontsize = 5))}) 


  pdf("fig3/3d_Top10_lung_heatmap_disease_complete.pdf",width=4,height=10)
  draw(p)
  dev.off()
```

# fig3e:TCR highlight Density (group:disease) 
```{r}
set.seed(200)
rcl.list <- row_order(p)
```

```{r}
TCRID_clus <- list()
meta_highlight <- list()
barcodes <- list()
TCR_cells <- list()
for(i in 1:length(rcl.list)){
TCRID_clus[[i]] <- rownames(mat1[rcl.list[[i]],])

meta_highlight[[i]] <- scRNAsubT8[,scRNAsubT8$TCR_ID %in%  TCRID_clus[[i]]]@meta.data
barcodes[[i]] <- rownames(meta_highlight[[i]])
TCR_cells[[i]] <- as.data.frame(cbind(as.character(meta_highlight[[i]][["TCR_ID"]]),barcodes[[i]]))}

meta <- scRNAsubT8@meta.data
for(i in 1:length(rcl.list)){
highlight_TCR <- TCR_cells[[i]]$V1[match(rownames(meta), TCR_cells[[i]]$V2)]
meta[ , ncol(meta) + 1] <- highlight_TCR
colnames(meta)[ncol(meta)] <- paste0("highlight_TCR_clus", i)}

meta <- meta %>% dplyr::select(starts_with("highlight_TCR_clus"))
md <- cbind(meta,Embeddings(scRNAsubT8[["umap"]]),scRNAsubT8[["sample_type"]])

getPalette = colorRampPalette(brewer.pal(10, "Paired"))
```

```{r}
plot.list <- list()
for (i in 1:length(rcl.list)){
  plot.list[[i]] = ggplot(md[!is.na(md[i]),], 
                          aes_string(x = "UMAP_1", y = "UMAP_2"))+
    geom_point(data = md[is.na(md[i]),], color = "lightgrey", size = 1)  + 
    geom_pointdensity(adjust= 0.5) +
    facet_wrap(~sample_type)+
    scale_color_gradientn(colors = rev( c("#581845", "#900C3F","#C70039","#FF5744","#FFC30F","#FBF9C1")))+
    guides(col = guide_legend(ncol = 1))+
    theme_cowplot()+
    theme_blank+
    theme(strip.background = element_blank(),
          strip.text = element_blank())+
    NoLegend()
  
    ggsave(paste("fig3/3e_DimDensity_TCR_disease_clus",i,".png",sep = ""),plot.list[[i]],width = 12, height = 4)
}
```

```{r}
plot.list <- list()
for (i in 1:length(rcl.list)){
  plot.list[[i]] = ggplot(md[!is.na(md[i]),], 
                          aes_string(x = "UMAP_1", y = "UMAP_2"))+
    geom_point(data = md[is.na(md[i]),], color = "lightgrey", size = 1)  + 
    geom_pointdensity(adjust= 0.5) +
    facet_wrap(~sample_type)+
    scale_color_gradientn(colors = rev( c("#581845", "#900C3F","#C70039","#FF5744","#FFC30F","#FBF9C1")))+
    theme(legend.title = element_text(size = 16), legend.text = element_text(size = 14))
}
l.list <- list()
for (i in 1:length(rcl.list)){
  l.list[[i]]=as_ggplot(get_legend(plot.list[[i]])) 
  ggsave(paste("fig3/3e_DimDensity_TCR_disease_clus",i,"_legend.pdf",sep = ""),l.list[[i]])}
```

# fig3f: pattern cell type stacked barplot
```{r}
TCRID_clus_group1 <- rownames(mat1[rcl.list[[1]],])
TCRID_clus_group2 <- rownames(mat1[rcl.list[[2]],])
TCRID_clus_group3 <- rownames(mat1[rcl.list[[3]],])

disease_pattern <- scRNAsubT8[,scRNAsubT8$TCR_ID %in% c(TCRID_clus_group1,TCRID_clus_group2,TCRID_clus_group3)]

df <- as.data.frame(cbind(c(TCRID_clus_group1,TCRID_clus_group2,TCRID_clus_group3),
                          rep(c("pattern 1a","pattern 1b","pattern 2"),
                              c(length(TCRID_clus_group1),
                                length(TCRID_clus_group2),
                                length(TCRID_clus_group3)))))

disease_pattern$pattern <- df$V2[match(disease_pattern$TCR_ID, df$V1)]
saveRDS(disease_pattern, "rds/T8_disease_TCRpattern.rds")
```

```{r}
clusters <- data.frame(table(disease_pattern$cell_type_main, disease_pattern$pattern))
colnames(clusters)[[2]]<-"pattern"
colnames(clusters)[[1]]<-"cell_type"


p =ggplot(clusters) + 
  geom_bar(aes(x =pattern, y= Freq, fill = cell_type),stat = "identity",width = 0.7)+ 
  labs(x=NULL,y = 'Number of Cells')+
  theme_cowplot()+
  scale_fill_manual(values=color_T8_main, limits = order_T8_main)+
  theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1,size = rel(1)),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14))+
  scale_y_continuous(expand = c(0,0))
ggsave("fig3/3f_pattern_celltype_stackedBar.pdf",p,width = 4, height = 5)
```

# fig3g: clonetype migration (group:disease) 
```{r}
disease <- scRNAsubT8[,scRNAsubT8$disease %in% "Pneumonia"]

metadata <- disease@meta.data
da.tcr.st1 <- metadata %>% 
  ungroup() %>% 
  mutate(Cell_Name=rownames(metadata), 
         clone.id = metadata$TCR_ID, 
         majorCluster=metadata$cell_type_main, 
         loc=metadata$sample_type) %>%
  dplyr::select(Cell_Name,clone.id,patient,majorCluster,loc,group,patient) %>% 
  as.data.frame()
da.tcr.st2 <- na.omit(da.tcr.st1)
```

```{r}
obj <- new("Startrac",da.tcr.st2,aid="T8cell")
obj <- calIndex(obj)
clus.data = obj@cluster.data

tic("pIndex")
obj <- pIndex(obj)
toc()

df.migr <- obj@pIndex.migr
df.migr <- df.migr[,-1]
rownames(df.migr)=df.migr[,1]
df.migr=df.migr[,-1]
df.migr <- as.data.frame(df.migr)

#gather()函数将宽格式转化为长格式
#df.migr为需要转的数据集名,migr为原数据列名的新列名,value为值的新列名
df.migr.2 <-gather(df.migr, migr, value, 1:3)
df.migr.2$majorCluster <- rep(rownames(df.migr),3)


df.migr.2$majorCluster<- factor(df.migr.2$majorCluster, levels =order_T8_main)
df.migr.2 <- subset(df.migr.2,df.migr.2$majorCluster != "T8_naive")
# no break
p = ggplot(df.migr.2, aes(x=majorCluster, y=value, fill=migr)) +
    geom_bar(stat="identity", position=position_dodge(), width=.8)+
    scale_y_continuous(expand=c(0,0))+
    coord_cartesian(ylim = c(0, 0.75))+
    theme_cowplot()+
    theme(axis.text.x = element_text(angle = 45,hjust =1, vjust = 1),
           axis.text = element_text(size = 16),
           axis.title = element_text(size = 16),
           legend.title = element_text(size = 16),
           legend.text = element_text(size = 14),
           legend.position = c(0.1,0.8),legend.direction = "vertical")+
    scale_fill_manual(values=color_migr)+
    xlab(NULL)+ylab("T8 TCR migration index")

ggsave("fig3/3g_migration_celltype.pdf",p,width = 5, height = 5)
```

# fig3h singleR schema
# singleR defined ctrl pattern
## ref and query data
```{r}
TCRID_clus_group1 <- rownames(mat1[rcl.list[[1]],])
TCRID_clus_group2 <- rownames(mat1[rcl.list[[2]],])
TCRID_clus_group3 <- rownames(mat1[rcl.list[[3]],])

lung_disease <- scRNAsubT8[,scRNAsubT8$sample_type %in% "Lung" & scRNAsubT8$disease %in% "Pneumonia"]
lung_disease <- lung_disease[,lung_disease$TCR_ID %in% c(TCRID_clus_group1,TCRID_clus_group2,TCRID_clus_group3)]

df <- as.data.frame(cbind(c(TCRID_clus_group1,TCRID_clus_group2,TCRID_clus_group3),
                          rep(c("pattern1a","pattern1b","pattern2"),
                              c(length(TCRID_clus_group1),
                                length(TCRID_clus_group2),
                                length(TCRID_clus_group3)))))

lung_disease$pattern <- df$V2[match(lung_disease$TCR_ID, df$V1)]
reference <-lung_disease
```

```{r}
lung_ctrl <- scRNAsubT8[,scRNAsubT8$sample_type %in% "Lung" & scRNAsubT8$disease %in% "Ctrl" &
                        !(scRNAsubT8$patient %in% "C2")]
lung_ctrl <- lung_ctrl[,lung_ctrl$freq_bchain_sample > 5]
query <- lung_ctrl
```

```{r}
reference_count=reference[["RNA"]]@counts
reference[['Index']] <- rownames(reference@meta.data)
pdata=reference@meta.data[,c("Index","pattern")]
rownames(pdata)=pdata$Index
pdata$Index=NULL
colnames(pdata)="pattern"
reference_SE <- SummarizedExperiment(assays=list(counts=reference_count),colData = pdata) 
reference_SE <- scuttle::logNormCounts(reference_SE)
```


```{r}
query_count=query[["RNA"]]@counts

common_gene <- intersect(rownames(query_count), rownames(reference_SE))
reference_SE <- reference_SE[common_gene,]
query_count <- query_count[common_gene,]

query_SE <- SummarizedExperiment(assays=list(counts=query_count))
query_SE <- scuttle::logNormCounts(query_SE)
```

## singleR
```{r}
singleR_res <- SingleR(test = query_SE, ref = reference_SE, labels = reference_SE$pattern)

anno_df <- as.data.frame(singleR_res$labels)
anno_df$Index <- rownames(singleR_res)
colnames(anno_df)[1] <- 'ref_label_from_ref'

query[['Index']] <- rownames(query@meta.data)
query$pattern <- anno_df$ref_label_from_ref[match(rownames(query@meta.data), anno_df$Index)]
```

## fig3i: singleR heatmap
```{r}
annoCol<-list(Labels=color_pattern)
p = plotScoreHeatmap(singleR_res,annotation_colors = annoCol)
ggsave("fig3/3i_singleR_heatmap.pdf",p,width = 5,height = 2)
```

## fig3j :singleR ctrl pattern
```{r}
ctrl_lung <- scRNAsubT8[,scRNAsubT8$disease %in% "Ctrl" & scRNAsubT8$sample_type %in% "Lung" & 
                        !(scRNAsubT8$patient %in% "C2")]
ctrl_lung[["pattern"]] <- NA
ctrl_lung$pattern <- anno_df$ref_label_from_ref[match(rownames(ctrl_lung@meta.data), anno_df$Index)]
```

```{r}
md <- cbind(ctrl_lung[["sample_type"]], ctrl_lung[["pattern"]],Embeddings(ctrl_lung[["umap"]]))
md$pattern <- factor(md$pattern, levels = order_pattern)

p = ggplot(md[!is.na(md$pattern),], aes(x = UMAP_1, y = UMAP_2, color = pattern))+
  geom_point(data = md[is.na(md$pattern),], color = "lightgrey", size = 1)  + 
  geom_point(shape = 16,size = 1.5,alpha = 1)  +
  guides(col = guide_legend(ncol = 1))+
  scale_color_manual(values =color_pattern)+
  theme_cowplot()+theme_blank + NoLegend()
ggsave("fig3/3j_singleR_DimPlot_ctrl_lung_origpattern.png",p,width = 5, height = 5)
```

```{r}
p = ggplot(md[!is.na(md$pattern),], aes(x = UMAP_1, y = UMAP_2, color = pattern))+
  geom_point(data = md[is.na(md$pattern),], color = "lightgrey", size = 1)  + 
  geom_point(shape = 16,size = 1.5,alpha = 1)  +
  guides(col = guide_legend(ncol = 1))+
  scale_color_manual(values =color_pattern)+
  theme_cowplot()+theme_blank
l=as_ggplot(get_legend(p)) 
ggsave('fig3/3j_singleR_DimPlot_ctrl_lung_origpattern_legend.pdf', l)
```

## fig3j: ctrl lung pattern bar
```{r}
clusters <- data.frame(table(ctrl_lung$pattern, ctrl_lung$sample_type))
colnames(clusters)[1:3]<- c("pattern","sample_type","Freq")

p = ggplot(clusters) + 
  geom_bar(aes(x =pattern, y= Freq, fill = pattern),
           position ="dodge",stat = "identity",width = 0.5) + 
  scale_fill_manual(values=color_pattern)+
  geom_text(data = clusters , aes(x = pattern, y=Freq, label=Freq),
            vjust= -0.5,fontface = "plain", family = "sans")+
  labs(x=NULL,y = 'Number of Cells')+
  theme_cowplot()+
  ylim(0,2500)+
  theme(axis.text.x = element_text(angle = 45,hjust =1, vjust = 1),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14))+
  NoLegend()
ggsave("fig3/3j_ctrllung_pattern_stackbar.pdf",p,width = 2.5, height = 4)
```
